From noise to knowledge: how randomness generates novel phenomena and information

### 2017-10-21 appendix/appendix.Rmd
WIP

```{r}
knitr::opts_chunk(message=FALSE)
```

@@

p <- list(r = .4, K = 1, sigma = .15, N = 500)

df %>% 
  filter(x > 0) %>% 
  group_by(t) %>%
  summarise(ave = mean(x), sd = sd(x)) %>%
  write_csv("inflation.csv")

Analytical predicted population size, as a fraction of $K$, is:

```{r}
(1 + sqrt(1 - 8 * p$sigma^2/p$K^2) )/ 2
```
Note this is independent of $r$, and increases with $\sigma/K$, the larger the noise $\sigma_g$ as a fraction of the carrying capacity $K$.  Population sizes can also be inflated relative to their deterministic equilibrium whenever the second derivative is positive (as in the lower equilibrium in the May model of alternative stable states, considered below.)  Stronger nonlinearities (in particular, a larger second derivative at the equilibrium) can drive larger inflationary effects.

```{r}
df %>% filter(t > 20, x > 0) %>%  summarize(mean(x))
```

df %>% 
  filter(x > 0) %>% 
  group_by(t) %>%
  summarise(ave = mean(x), sd = sd(x)) %>%
  gather(moment, value, -t) %>%
  ggplot(aes(t,value, col=moment)) + geom_line()

@@

@@

Needs a model with non-zero alternate state!

@@

tibble(t = seq_along(cmodel$x), x = cmodel$x) %>% write_csv("tipping.csv")

>>>>

```{r}
knitr::opts_chunk$set(message=FALSE)
```

@@

@@

Analytical predicted population size, as a fraction of $K$, is $\tfrac{1}{2} + \tfrac{1}{2} \sqrt{1- 8\sigma^2/K^2}$.  Note this is independent of $r$, and increases with $\sigma/K$, the larger the noise $\sigma_g$ as a fraction of the carrying capacity $K$.  Population sizes can also be inflated relative to their deterministic equilibrium whenever the second derivative is positive (as in the lower equilibrium in the May model of alternative stable states, considered below.)  Stronger nonlinearities (in particular, a larger second derivative at the equilibrium) can drive larger inflationary effects.

```{r}
p <- list(r = .4, K = 1, sigma = .15, N = 500)
equib <- (1 + sqrt(1 - 8 * p$sigma^2/p$K^2) )/ 2

df %>% 
  filter(x > 0) %>% 
  group_by(t) %>%
  summarise(ave = mean(x), sd = sd(x)) %>%
  mutate(equib = equib) %>%
  write_csv("inflation.csv")

read_csv("inflation.csv") %>% 
  ggplot(aes(t,value, col=moment)) + geom_line() + 
  geom_hline(aes(yintercept = equib))

----

@@

@@

tibble(t = seq_along(cmodel$x), x = cmodel$x) %>% 
  mutate(var = RcppRoll::roll_var(x, n = 5)) %>%
  write_csv("tipping.csv")

### 2017-10-21 paper.Rmd
WIP

@@

```{r}
read_csv("appendix/quasicycles.csv") %>%
  gather(species, pop, -t, -sigma) %>%
  ggplot(aes(t, pop, col=species)) + geom_line() + 
  facet_wrap(~sigma) + scale_color_ipsum() + 
  labs(x="Time", 
       y="Population density",
       title="Quasi-cycles created by stochastic resonance",
       subtitle="Left panel: neglible  noise results only in damped oscillations.
                 Right panel: small noise creates sustained oscillations.")
```

- Oscillations (parametric oscillator, stochastic resonance)

@@

@@

- population inflation / deflation

>>>>

<!--

We hope that these papers will combine a mix of synthetic reviews of the current state of an area, as well as insights into how the field has developed during the last twenty years, discussion of current important questions, and developments and insights that would significantly advance the field. Although certainly not necessary, if there are any past Ecology Letters papers that have had a profound impact on the development of an area – be it beneficial or not – a brief discussion of that impact would be appropriate. We also hope that articles will be written in a thoughtful style, but also provocative if appropriate.

Key questions

Key advances

Shift from noise as nuiance to noise as opportunity
Stochastic phenomena outside of theory -- early warning signals

Need for greater tanslation

Population-level vs ABM vs IBM

Discuss out of scope:  
 - behavior
 - evolutionary dynamics examples??

-->

@@

$$x_{t+1} = x_t + x_t r \left(1 - \frac{x_t}{K}\right) - b x_t  y_t + \epsilon_t$$
$$y_{t+1} = y_t + c  x_t  y_t - d  y_t+ \xi_t $$

```{r}
read_csv("appendix/quasicycles.csv") %>%
  gather(species, pop, -t, -sigma) %>%
  ggplot(aes(t, pop, col=species)) + geom_line() + 
  facet_wrap(~sigma) + scale_color_ipsum() + 
  labs(x="Time", 
       y="Population density",
       title="Quasi-cycles created by stochastic resonance",
       caption="Left panel: neglible  noise results only in damped oscillations.\nRight panel: small noise creates sustained oscillations.")
```

- Disease models

- Power spectrum

@@

**Stochastic switching**

Stochastic fluctuations can also drive spontaneous shifts between alternative stable states.

```{r}
read_csv("appendix/noisy_switch.csv")  %>% ggplot(aes(t,x)) + 
  geom_line() + 
  labs(title="Stochastic switching", 
       subtitle = "Noise drives transitions between alternative stable states")
```

- transition probabilities, waiting times

@@

Stochasticity can

```{r}
read_csv("appendix/inflation.csv") %>% 
  ggplot(aes(t,value, col=moment)) + geom_line() + 
  geom_hline(aes(yintercept = equib))
```

**Bet hedging**

**Further examples**
